---
title: "HW 2: Visualization, Base R"
subtitle: "Data 180, Professor Bilen"
author: 
  name: "Hoa Ly"
  email: "lyh@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`' # no need to change date since this line grabs today's date
output: 
  html_document
---

Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 26 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:

```{r}
# Your code goes here
```

Make sure to insert a chunk for each (sub)section of the assignment.

**Theme of the assignment**: Let's practice visualization using base R. Load in the `Cars93` dataset by running the following code chunk:

```{r}
library(MASS)
data(Cars93)
```

The library function will load and attach add-on packages that have been previously installed on your machine. This will give you access to the functions and datasets available in the package. In general, you only need to install a package once (although it may require updating from time to time) but you need to access it using the library command every time you start a new session in R (and want the functionality).

The questions below use the `Cars93` data frame as a source of data. This data frame is part of the MASS package and becomes available after loading it. You can get a view of the top of the data frame by using the head command. For more information, type `?Cars93`. Let's load `Cars93` in,

```{r}
head(Cars93)
```

# Base R

## Task 1

a)  Create a Frequency table and Relative Frequency table (latter uses percentages) for the variable `Type`. Make sure to round all digits to 3 digits.

```{r}
table(Cars93$Type)
round(table(Cars93$Type)/nrow(Cars93),3)
```

b)  Create a frequency bar graph for the variable `Type`. You can have count on the y axis.

```{r}
barplot(main = "Type Frequency bar graph", table(Cars93$Type), ylab = "Count", xlab = "Type", col = "yellow", cex.lab = 1.2, cex.names = 1.2)
```

-   Your axes should be labeled and legible
-   The bars should be of different colors
-   The plot should include a title using the main argument
-   The names, labels, and axes sizes (cex parameter) should be set to 1.2

c)  Create a pie chart of the variable `Type`.

```{r}
pie(main = "Type pie chart", table(Cars93$Type), cex.lab = 1.2)
```

-   The plot should include a title using the main argument.
-   The names, labels, and axes sizes (cex parameter) should be set to 1.2

d)  What is the mode of the variable `Type`? Comment. 

The mode of the variable 'Type' is Midsize

## Task 2

a)  Create a histogram of the variable `Horsepower`.

```{r}
hist(Cars93$Horsepower,main = 'Horsepower histogram', xlab = 'Horsepower', col = "yellow", cex.lab = 1.2, cex.axis = 1.2)
```

-   Your axes should be labeled
-   The hist body should have a color (your choice, not too dark to obscure lines)
-   The plot should include a title using the main argument
-   The names, labels, and axes sizes should be set to 1.2

b)  Create a second histogram with breaks at 25 HP increments instead of 50.

```{r}
hist(Cars93$Horsepower, breaks = seq(50,300,25), main = 'Horsepower histogram', xlab = 'Horsepower', col = "yellow", cex.lab = 1.2, cex.axis = 1.2)
```

-   Use the breaks argument to define the classes
-   Other setting should be as in part (a)

c)  Which histogram more closely follows or rule of thumb for the suggested number of classes? Explain.

```{r}

```

## Task 3

a)  What is the type of variable `Passengers`? (Categorical nominal/ordinal or Numerical discrete/continuous?) What happens if we were to make a histogram for `Passengers`? What is the problem with this graph?

```{r}
'The type of the variable "Passengers" is Numerical discrete. Since we can only display continuous value with a histogram, it would not be possible to display a discrete value with a histogram"
```

b)  Correct the problem in part (a) by creating a more appropriate visualization. Make sure to label your axes, place a title, choose a custom color as in the previous graphs. (Bonus: use factor() with levels option to display x axis values with zero observations on the plot. Hint: when you have discrete values, it might make sense to treat each value as a category.)

```{r}
barplot(table(Cars93$Passengers), main = 'Passengers seat chart',col='yellow',ylab='frequency',xlab='Number of passengers',cex.lab=1.2,cex.names=1.2)
```

## Task 4

a)  Make a histogram of the variable `Length` (the default breaks are fine). Other aspects of the plot should be as in part (a) in Task 2.

```{r}
hist(Cars93$Length, main = 'Length histogram', xlab = 'Length', col = "yellow", cex.lab = 1.2, cex.axis = 1.2)
```

b)  What is the shape of the distribution? the distribution is symmetric.

c)  Find the mean and median of the variable `Length.` Do these answers meet your expectations from part (b)?

```{r}
mean(Cars93$Length)
median(Cars93$Length)
'These answers does meet my expectations from part (b)'
```

You are done! 🏁 Don't forget to commit and push your .Rmd file to your Github repository before the due date.
